---
title: "R Notebook"
output: html_notebook
---

```{r setup}
library(tidyverse)
library(sf)
library(rnaturalearth)
```


```{r}
turbine <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-10-27/wind-turbine.csv')
```



Let's get started. The data is on Canadian wind turbines. I haven't mapped data in a while, so that's something I'd like to do. I don't know much about Canada's geography, so that'll take some getting used to. Still, shouldn't be hard to get a basic plot. Where to get the Canada map data in SF format? Probably can get something from `rnaturalearth`.

```{r}
canada <- rnaturalearth::ne_countries(country = "canada", 
                                      scale = "medium") %>% 
  as("sf")

plot(st_geometry(canada), graticule = TRUE, axes = TRUE)
```

Okay, it's not the best map; it lacks political borders. Still, it's a good start. Let's merge some stuff in. To do that, I need to convert 'turbine' to an `sf` object. 

```{r}
names(turbine)
```

`longitude` and `latitude` are points, so I need to convert them to a list of sf_POINT objects. 

```{r}
longlat <- vector(mode = 'list')

for (i in seq(length(long))) {
  longlat[[i]] <- st_point(
    c(turbine$longitude[i], turbine$latitude[i])
  )
}

turbine_sf <- turbine %>% 
  mutate(longlat = st_sfc(longlat)) %>% 
  st_sf(crs = 4326)

rm(list = c('longlat', 'i'))

```

Let's see how it looks. 

```{r}
plot(st_geometry(canada))
plot(st_geometry(turbine_sf), add = TRUE)
```

Nice. Let's move to `ggplot2` and fix up the margins, make it look nice. 

```{r}
ggplot() + 
  geom_sf(data = canada) + 
  geom_sf(data = turbine_sf)
```

Looking nice, but the projection could be better. It's so high, is all, and would be better served with a rounder projection. I don't know much about these projections. 

```{r}
st_is_longlat(turbine_sf)
st_is_longlat(canada)
```

Both of these are unprojected except for their default projections. I can try to set my own re-projection of Canada, even though I'm not sure what's required. 

```{r}
canada_t <- canada %>% 
  st_transform(3573)
turbine_sf_t <- turbine_sf %>% 
  st_transform(3573)

ggplot() +
  geom_sf(data = canada_t, fill = 'black') + 
  geom_sf(data = turbine_sf_t, color = 'red') + 
  theme_minimal()
```

Okay, something I should've known but didn't is that when you have a simple point geometry, you can use `sf::st_as_sf(data, coords = c("x","y"))` to convert to an SF object in one line. The more you know. Let's look at this data over time. 



Some notes from my work in progress for this week: 

```{r}
# Before I go animate every point, let's do some tests
df <- data.frame(
  x = c(1, 2, 3, 4, 5, 6, 7),
  y = c(5, 3, 5, 7, 9, 0, 9),
  xy_group = c(2010, 2010, 2015, 2015, 2015, 2015, 2019)
)

p1 <- df %>% 
  ggplot() + 
  geom_point(aes(x = x, y = y, group = xy_group)) + 
  transition_states(xy_group) + 
  exit_shrink()
animate(
    p1, nframes = 20
  )


# Trickier is to make the data points permanent when they appear. This involves
# changing the data. I need to fill in the data, each observation with every
# year. Then I can give a 1 or a 0 if the turbine has been built already.

# Candidate functions: tidyr::complete (which is basically tidyr::expand
#                      expand.grid

# First, what exactly do I want the dataset to look like? 
df <- data.frame(
  x = c(1, 1, 1, 2, 2, 2, 3, 3, 4, 4, 5, 5, 6, 6, 7),
  y = c(5, 5, 5, 3, 3, 3, 5, 5, 7, 7, 9, 9, 0, 0, 9),
  xy_group = c(2010, 2015, 2019, 2010, 2015, 2019, 2015, 2019, 2015, 2019, 2015, 2019, 2015, 2019, 2019)
)

# This won't solve the transition problem, though. Tricky. Aha, the answer is
# transition_reveal

df <- data.frame(
  x = c(1, 2, 3, 4, 5, 6, 7),
  y = c(5, 3, 5, 7, 9, 0, 9),
  xy_group = c(2010, 2010, 2015, 2015, 2015, 2015, 2019)
)

p1 <- df %>% 
  ggplot() + 
  geom_point(aes(x = x, y = y, group = seq_along(xy_group))) + 
  transition_reveal(seq_along(xy_group))
animate(
  p1, nframes = 21
)

# In the case of time, gganimate has several options
p1 <- df %>% 
  ggplot() + 
  geom_point(aes(x = x, y = y, group = xy_group)) + 
  transition_time(xy_group)

animate(
  p1, nframes = 21
)

# This is ideal, I think
p1 <- df %>% 
  ggplot() + 
  geom_point(aes(x = x, y = y), size = 4) + 
  transition_events(start = xy_group,
                    end = xy_group,
                    enter_length = 0,
                    exit_length = 5) + 
  exit_shrink()

animate(
  p1, nframes = 50
)
```




# Number of turbines within 50 miles as color
The key here is st_within, or something similar. I know I worked with these on the plant projects, so I can check there if I forget how it all works. 

The binary operations all return a sparse matrix by default. I need to simply get the number of other points within a certain distance. 

```{r}
distance <- st_distance(turbine_t)
```

This is what I needed, but it's large (6698x6698, 358 MB). That's okay, I suppose. Let me use a small version for now, so I can troubleshoot things. 

```{r}
turbine_subset <- turbine_t %>% 
  dplyr::slice_head(n = 10)

distance <- st_distance(turbine_subset)

n_within_800m <- apply(distance, 2, function(x) sum(x < 800) -1)
n_within_800m

turbine_subset %>% 
  mutate(
    n_within_800m = apply(distance, 2, function(x) sum(x < 800) -1)
  ) %>% 
  ggplot() + 
  geom_sf(aes(geometry = geometry, color = n_within_800m))
```

Ah cool, I'm happy this works. Let's see if we can put it all together, then.


```{r}
distance <- st_distance(turbine_t)

turbine_t %>% 
  mutate(
    n_within_10000m = apply(distance, 2, function(x) sum(x < 10000) -1)
  ) %>% 
  ggplot() + 
  geom_sf(aes(geometry = geometry, color = n_within_10000m),
          size = 1) + 
  scale_color_gradient(low = '#000000', high = '#ffffff') + 
  theme(
    plot.background = element_rect(fill = 'black'),
    panel.background = element_rect(fill = 'black')
  )
```


