---
title: "Government Spending on Kids"
output: html_notebook
---

```{r, include=FALSE}
library(tidyverse)
library(janitor)
```

```{r}
kids <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-15/kids.csv')

kids_raw <- pivot_wider(data = kids, id_cols = c(state, year), names_from = variable, values_from = raw) %>%
  janitor::clean_names()
kids_inf_adj <- pivot_wider(data = kids, id_cols = c(state, year), names_from = variable, values_from = inf_adj) %>%
  janitor::clean_names()
kids_inf_adj_perchild <- pivot_wider(data = kids, id_cols = c(state, year), names_from = variable, values_from = inf_adj_perchild) %>% 
  janitor::clean_names()

skimr::skim(kids_raw)
```

The data is very long because they include several measures of the same variables. I widened each one to make it easier to understand them. Here is the raw data point data, skimmed. There are many variables, none of which are listed on the data dictionary on the Github site; you have to go to the package source. (Technically, this data comes from Urban Institute, repackaged by Joshua Rosenberg in his `tidykids` package.) The full codebook can be found [here](https://jrosen48.github.io/tidykids/articles/tidykids-codebook.html). To put it mildly, the data is dense. 


- `state` refers to the US, including DC
- `year` Year
- `pk12ed` Public spending on elementary and secondary education by state and year, in $1,000s
- `highered` Public spending on higher education by state and year, in $1,000s
- `edsubs` Public spending on education subsidies, including tuition and scholarships, by state and year, in $1,000s
- `edservs` Public spending on education special services by state and year, in $1,000s
- `pell` 	Public spending on Pell Grants by state and year, in $1,000s
- `head_start_priv` Federal spending on Head Start awarded to private grantees by state and year, in $1,000s
- `tan_fbasic` Public spending on TANF cash assistance payments by state and year, in $1,000s
- `othercashserv` Public spending on other cash assistance payments and social services by state and year, in $1,000s
- `snap` Public spending on SNAP benefit payments that go to children by state and year, in $1,000s
- `socsec` Public spending on Social Security payments that go to children by state and year, in $1,000s
- `fed_ssi` Federal spending on SSI payments that go to children by state and year, in $1,000s
- `fed_eitc` Federal spending on EITC, including refundable and non-refundable portions, by state and year, in $1,000s
- `ctc` Public spending on the Child Tax Credit, including refundable and non-refundable portions, by state and year, in $1,000s
- `add_cc` Public spending on the Additional Child Tax credit, including only refundable portion, by state and year, in $1,000s
- `state_eitc` Total state spending on EITC by state and year, in $1,000s
- `unemp` Public spending on Unemployment Compensation benefit payments by state and year, in $1,000s
- `wcomp` Public spending on Workers Compensation benefit payments by state and year, in $1,000s
- `medicaid_chip` Public spending on Medicaid (for children and youth <21) and CHIP by state and year, in $1,000s. This variable has a very long description; use caution.
- `pubhealth` Public spending on public health efforts by state and year, in $1,000s
- `other_health` Public spending on health vendor payments and public hospitals, excluding Medicaid, by state and year, in $1,000s
- `hcd` Public spending on housing and community development by state and year, in $1,000s
- `lib` Public spending on libraries by state and year, in $1,000s
- `parkrec` Public spending on parks and recreation by state and year, in $1,000s


Further important definitions: 

- __TANF__ Temporary Assistance for Needy Families
- __SNAP__ Supplemental Nutrition Assistance Program
- __EITC__ Earned Income Tax Credit

State and Federal spending is a complex topic that I don't have much experience with, so a simple graphic will go furthest this week. I've never made one of those square composition charts. Let me see if I can find one. Ah a `treemap`, that's it. Alternatively I could do a marimekko chart, with variable sized bars. This would be great for visualizing spending per capita multiplied by the number of persons in each state. (This would be able to accommodate only one variable at a time, however.) Then, spending per capita on the y-axis, population defines the width the bar. Let me start with the marimekko chart and move to the treemap, which is much harder to customize on short notice. 

```{r}
# test
df <- tibble(
  x = c(1, .3, .5, .4, .2),
  y = c(5, 4, 3, 2, 1),
  title = c('one','two','three','four','five')
)

df %>% 
  ggplot() + 
  geom_col(aes(x = title, y = y, width = x))

```

Well, I had hoped that this would be a cinch, but it seems I will have to define the rectangles myself and use `geom_rect()`. This shouldn't be too hard, but it's not great. Let me try with the example dataset above. I need to generate the x max/min and y max/min. 

- `ymin` is always zero. 
- `ymax` is simply the y value
- `xmin` and `xmax` need processing. First, I need center positions of each bar; second, I need to generate the extremities of the bars. This needs to be done so that there is a constant amount of space between each bar. I'll start with bars that touch and work inwards from there.

`xcenter` can be defined with the `cumsum` function. Unless I'm mistaken, you can take the average of the cumulative sum and the lagged cumulative sum to generate a center point; but then, with a cumulative sum you can probably outright generate the extremities. 

```{r}
df <- tibble(
  x = c(1, 3, 5, 4, 2),
  y = c(5, 4, 3, 2, 1),
  title = c('one','two','three','four','five')
)

df %>% 
  mutate(
    ymin = 0,
    ymax = y,
    xmax = cumsum(x),
    xmin = lag(xmax, default = 0)
  ) %>% 
  ggplot() + 
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = ymax, fill = title))


```

I'll introduce space by shifting the bars over by a fixed amount. This way, the bars retain their constant width. To shift the bars over by a constant amount, I need to add progressively more to each boundary point. 

```{r}
df <- tibble(
  x = c(1, 3, 5, 4, 2),
  y = c(5, 4, 3, 2, 1),
  title = c('one','two','three','four','five')
) %>% 
  mutate(
    ymin = 0,
    ymax = y,
    xmax = cumsum(x),
    xmin = lag(xmax, default = 0),
    xminplus = xmin + 0.25*(row_number() - 1),
    xmaxplus = xmax + 0.25*(row_number() - 1),
    xcenter = (xminplus + xmaxplus) / 2
  )

df %>% 
  ggplot() + 
  geom_rect(aes(xmin = xminplus, xmax = xmaxplus, ymin = 0, ymax = ymax, fill = title))
```

That looks nice. I'm going to have problems with the x-axis no matter what, though, because the reader will have to do arithmetic to find the width of the bar. I have to be creative, instead. On the [mekko graphics](https://www.mekkographics.com/why-use-a-marimekko-chart/) website, they write the width of the bar at the top of the bar. That would look like this: 

```{r}
df_text <- df %>% 
  select(ymax, xcenter, title) %>% 
  mutate(
    xwidth = paste0(title, ": ", c('1','3','5','4','2')),
    xwidth_y = ymax + 0.25
  )


df %>% 
  ggplot() +
  geom_rect(aes(xmin = xminplus, xmax = xmaxplus, ymin = 0, ymax = ymax, fill = title)) + 
  geom_text(data = df_text, aes(x = xcenter, y = xwidth_y, label = xwidth))
```

That will have to do. For a theme, I'd like to use something simple and slightly technical looking. Let's do it with some real data. I need to (a) choose a variable and (b) merge in population data. Alternatively, I could add to the complexity by including an array of variables. I just realized I can derive the number of children. 


```{r}
ggkids <- kids %>% 
  filter(year == 2016, variable == "PK12ed") %>% 
  arrange(desc(inf_adj_perchild)) %>% 
  mutate(
    n_kids = inf_adj / inf_adj_perchild,
    ymin = 0,
    ymax = inf_adj_perchild,
    xmax = cumsum(n_kids),
    xmin = lag(xmax, default = 0),
    xcenter = (xmax + xmin) / 2
  )

ggkids_text <- ggkids %>% 
  select(ymax, xcenter, state, n_kids) %>% 
  mutate(
    xwidth = paste0(state, ": ", as.character(round(n_kids / 1000), n = 2), "k"),
    xwidth_y = ymax - 0.2
  )

ggkids %>% 
  ggplot() + 
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = ymax, fill = n_kids)) + 
  geom_text(data = ggkids_text, aes(x = xcenter, y = xwidth_y, label = xwidth),
            angle = 90, hjust = 1, color = 'white')
```

This might look better if I rotate the axes, so the text is read normally and the bars are going to the right. That means doing another long graphic, but I'm more or less okay with that. 

```{r}
ggkids <- kids %>% 
  filter(year == 2016, variable == "PK12ed") %>% 
  arrange(desc(inf_adj_perchild)) %>% 
  mutate(
    n_kids = inf_adj / inf_adj_perchild,
    xmin = 0,
    xmax = inf_adj_perchild,
    ymax = cumsum(n_kids),
    ymin = lag(ymax, default = 0),
    ycenter = (ymax + ymin) / 2
  )

ggkids_text <- ggkids %>% 
  select(xmax, ycenter, state, n_kids) %>% 
  mutate(
    ywidth = paste0(state, " (", as.character(round(n_kids / 1000), n = 2), "k)"),
    ywidth_x = xmax - 0.2
  )

ggkids %>% 
  ggplot() + 
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = n_kids)) + 
  geom_text(data = ggkids_text, aes(y = ycenter, x = ywidth_x, label = ywidth),
            angle = 0, hjust = 1, color = 'white') + 
  scale_x_continuous(name = "Public spending on elementary and secondary education by state and year, in $1,000s")
```

I think this is plenty for an interesting graphic. I can figure out the colors of the bars and the theme in the R file. 

I've lost interest in the verticality of this. It seems universal to use a long x-axis instead, so that's what I'm going to do. 