---
title: "Taylor Swift"
output: html_notebook
---

This week's data: Taylor Swift. 

```{r,include=FALSE}
library(tidyverse)
```

### Get the data
```{r}
beyonce_lyrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/beyonce_lyrics.csv')
taylor_swift_lyrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/taylor_swift_lyrics.csv')
sales <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/sales.csv')
charts <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/charts.csv')
```

It looks like there are four datasets. (1) Beyonce lyrics, (2) Taylor Swift lyrics, (3) sales, and (4) charts.

I don't have any particular inspiration for this graphic. I don't like sentiment analysis of lyrics because the sentiment is heavily weighted towards songs with more (and more repetitive) lyrics. There are a host of problems with sentiment analysis that I don't know how to solve, too. What about negations? Does 'not happy' count as positive, negative, or neutral? I'm just not good enough at natural language processing yet. 

So, barring that I have to stick with a simpler linguistic analysis, or something that integrates other data. Let's get a feeling for the other data sets. 

```{r}
charts %>% 
  skimr::skim()
```

So, these are all characters right now, but I think chart_position is just a normal ranking. Let me see if I can clean it up at all. Convert date variables to date-time format. For ranking, replace "—" with "". 


```{r}
charts_clean <- charts %>% 
  mutate(
    chart_position = as.integer(gsub("—", "", chart_position)),
    released_date = gsub(" \\(US\\)\\[51\\]", "", released),
    released_date = gsub("June 23, 2003 \\(UK\\)\\[39\\]", 
                    "June 23, 2003", released_date),
    released_date = lubridate::mdy(released_date)
  ) %>% 
  select(artist, title, released, released_date, everything())
```

The unit of observation here is the artist/album/country. How can I visualize this, then? See notebook, first page. 

```{r}
charts_clean %>% 
  filter(artist == "Taylor Swift") %>% 
  ggplot() + 
  geom_col(aes(x = reorder(title,released_date), y = chart_position, fill = reorder(title, released_date))) + 
  facet_wrap(vars(chart)) + 
  scale_y_reverse() + 
  theme_void() + 
  theme(
    panel.grid.major.y = element_line(color = grey(0.9), linetype = 'solid')
  )
```

I really like this actually. I think this might be a good rough design. There's one other type of graphic I want to explore, having to do with concentric circles. But that graphic is better suited to segmented populations where you want to show each data point inside a circle. It's also not clear how I would put each data point in its circle... Probably with a polar coordinate system, then it shouldn't be too hard. Okay, I won't use it, but for the experiment I'll generate a usable dataset. 

Alternatively, I could put each point into a circle where the size of the circle is mapped to the population of the country. But what points would go in each circle? Agh. But it's an interesting idea. I want to explore it further. 


```{r}
df <- data.frame(
  x = runif(n = 50, min = 0, max = 1),
  y = runif(n = 50, min = 1, max = 10)
) %>% 
  mutate(
    group = case_when(
      y <= 6 ~ 'one',
      y > 6 & y <= 9 ~ 'two',
      y > 9 ~ 'three'
      ),
    group = factor(group)
    )

df %>% 
  ggplot() + geom_point(aes(x = x, y=y, color = group)) + 
  coord_polar() + 
  scale_y_continuous(breaks = c(6, 9, 10)) + 
  theme_void() + 
  theme(
    panel.grid.major.y = element_line(color = grey(0.8), linetype = 'solid')
  )
  
```

