---
title: "Great American Beer Festival"
output: html_notebook
---

```{r setup}
library(tidyverse)
library(skimr)
library(janitor)
```


```{r}
beer <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-10-20/beer_awards.csv')

skimr::skim(beer)
```

Simple dataset this week, which is good because I'm getting a late start on it. Beer we go. 

Five-thousand observations, each observation is a beer/year. Variables include medal (gold, silver, bronze), name, brewery, city, state, and category. There are 515 categories of beer, which is interesting. Which categories win the most? The least? 

```{r}
beer %>% 
  mutate(
    gold = medal == "Gold",
    silver = medal == "Silver",
    bronze = medal == "Bronze"
  ) %>% 
  group_by(category) %>% 
  summarize(
    n_gold = sum(gold),
    n_silver = sum(silver),
    n_bronze = sum(bronze),
    n_medals = sum(gold + silver + bronze)
  ) %>%
  arrange(n_medals)
    
```

That wasn't as interesting as I hoped it would be. Turns out its methodological nonsense. I wish they included taste profiles with these like they did with the coffee rankings. Anyway, I don't have an immediate idea of what visual I'd like to create this week. I'm thinking something black and white, like that one person on Twitter. Also, [this](https://github.com/gkaramanis/tidytuesday/blob/master/2020-week08/food-consumption.R) is a great bit a code from Georgios Karamanis. Pretty unbelievable actually, I'm really impressed. 

So, B/W with a red highlight is probably the direction I'm going to go. That could highlight California, but I also might track changes over time with a long bump chart. Let's look at that. 

How will I rank and group? Maybe not a bump chart, but cumulative medals over time for each state. That would be a step chart. 

```{r}
beer %>% 
  mutate(
    gold = medal == "Gold",
    silver = medal == "Silver",
    bronze = medal == "Bronze"
  ) %>%
  group_by(year, state) %>% 
  summarize(
    n_medals = sum(gold + silver + bronze)
  ) %>% 
  group_by(state) %>% 
  mutate(
    cum_medals = cumsum(n_medals),
    pa = state == "WI"
  ) %>% 
  # filter(year == 2020) %>% 
  # arrange(desc(cum_medals))
  ggplot() + 
  geom_step(aes(x = year, y = cum_medals, group = state, color = pa)) +
  scale_color_manual(values = c(grey(0.6),'black')) + 
  theme_void()
```

Well, it's not as good as the Navratilova graphic from last year, but I still like it. Let's run with it. Which state is the one that's flagging at the end? Ah, Wisconsin. Poor Wisconsin. Is there a possible reason for their loss of momentum? Impossible to say with this dataset, of course. Anyway, let's start finishing this graphic. 

